Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime \thanks{ This work has been supported by the Italian PRIN Project COMMON-WEARS (2020HCWWLP) and the EU/MUR FSE REACT-EU PON R\&I 2014-2020. }

### 2023-05-21 paper-2023-acsos-pulverization.tex
Minors

\begin{lstlisting}[
    float=ht,
    numbers=left,
    language=pulv,
    caption={
        Example of the usage of the domain-specific language for the runtime configuration.
        The example defines the runtime configuration of the system, specifying the initial deployment and the reconfiguration rules.
    },
    label=lst:runtime-configuration,
    numbers=left,
    escapechar=\%
]
// Reconfiguration events %\label{code:reconfigev:start}%
expect fun cpuLoad(): Flow<Double>
expect fun batteryLevel(): Flow<Double>
object HighLoad : ReconfigurationEvent<Double>() {
  override val predicate = { it > 0.90 }
  override val events = cpuLoad()
}
object LowBattery : ReconfigurationEvent<Double>() {
  override val predicate = { it < 0.20 }
  override val events = batteryLevel()
}%\label{code:reconfigev:end}%
// Available hosts %\label{code:hosts:start}%
object Smartphone : Host {
  override val hostname = "smartphone"
  override val capabilities = setOf(EmbeddedDevice)
}
object Server : Host {
  override val hostname = "amazon-aws"
  override val capabilities=setOf(HighCPU,LowLatencyComm)
}%\label{code:hosts:end}%
// Runtime initial setup and runtime reconfiguration rules %\label{code:infra:start}%
val infrastructure = setOf(Smartphone, Server)
val conf = // see %\crefe{lst:system-configuration}%
pulverizationRuntime(conf, "iot-sensor", infrastructure) {
  DeviceBehaviour() startsOn Server%\label{code:mapping:start}%
  DeviceCommunication() startsOn Server
  DeviceSensors() startsOn Smartphone
  DeviceActuators() startsOn Smartphone%\label{code:mapping:end}%
  reconfigurationRules {%\label{code:reconfig:start}%
    onDevice {
      HighLoad reconfigures {Behaviour movesTo Smartphone}
      LowBattery reconfigures {Behaviour movesTo Server}
    }
  }%\label{code:reconfig:end}%
}%\label{code:infra:end}%
pulverizationRuntime(conf, "control-center", infrastructure) {
  /* ... */
}
\end{lstlisting}

@@

The runtime bases its operation on three main components:
the \texttt{Spawner}, which is responsible for executing the components;
the \texttt{UnitReconfigurator}, which is responsible for reconfiguring the \emph{deployment unit}
according to the \texttt{ReconfigurationEvent}
and the \texttt{Communicator}s,
which are used to establish the components' intra-communication.
%
To start, the \texttt{PulverisationRuntime} needs to know the name of the \emph{logical device},
the \emph{host} on which it is running, and the configurations produced by the two \ac{dsl}s.
%
At the start of the system, the runtime,
based on the provided configuration,
spawns all the components that must start on the \emph{deployment unit}.
%
Similarly, the \texttt{UnitReconfigurator} starts observing all the configured \texttt{ReconfigurationEvent}s
and registers the listeners for incoming reconfigurations.
%
Then,
each spawned component starts
executing its logic and sharing data with other components via the \texttt{Communicator},
which is responsible for the components' intra-communication by leveraging the communication protocol.
%
Finally,
when a reconfiguration occurs, the \texttt{UnitReconfigurator} considers the new deployment configuration and,
by interacting with the \texttt{Spawner},
starts or stops components in order to achieve the new setup.

>>>>

\begin{lstlisting}[
    float=ht,
    numbers=left,
    language=pulv,
    caption={
        Example of the usage of the domain-specific language for the runtime configuration.
        The example defines the runtime configuration of the system, specifying the initial deployment and the reconfiguration rules.
    },
    label=lst:runtime-configuration,
    numbers=left,
    escapechar=\%
]
// Reconfiguration events %\label{code:reconfigev:start}%
expect fun cpuLoad(): Flow<Double>
expect fun batteryLevel(): Flow<Double>
object HighLoad : ReconfigurationEvent<Double>() {
  override val predicate = { it > 0.90 }
  override val events = cpuLoad()
}
object LowBattery : ReconfigurationEvent<Double>() {
  override val predicate = { it < 0.20 }
  override val events = batteryLevel()
}%\label{code:reconfigev:end}%
// Available hosts %\label{code:hosts:start}%
object Smartphone : Host {
  override val hostname = "smartphone"
  override val capabilities = setOf(EmbeddedDevice)
}
object Server : Host {
  override val hostname = "amazon-aws"
  override val capabilities=setOf(HighCPU,LowLatencyComm)
}%\label{code:hosts:end}%
// Runtime initial setup and runtime reconfiguration rules %\label{code:infra:start}%
val infrastructure = setOf(Smartphone, Server)
val conf = ... // see %\crefe{lst:system-configuration}%
pulverizationRuntime(conf, "iot-sensor", infrastructure) {
  DeviceBehaviour() startsOn Server%\label{code:mapping:start}%
  DeviceCommunication() startsOn Server
  DeviceSensors() startsOn Smartphone
  DeviceActuators() startsOn Smartphone%\label{code:mapping:end}%
  reconfigurationRules {%\label{code:reconfig:start}%
    onDevice {
      HighLoad reconfigures {Behaviour movesTo Smartphone}
      LowBattery reconfigures {Behaviour movesTo Server}
    }
  }%\label{code:reconfig:end}%
}%\label{code:infra:end}%
pulverizationRuntime(conf, "control-center", infrastructure) {
  /* ... */
}
\end{lstlisting}

@@

The runtime bases its operation on three main components:
the \texttt{Spawner}, which is responsible for executing the components;
the \texttt{UnitReconfigurator}, which is responsible for reconfiguring the \emph{deployment unit}
according to the \texttt{ReconfigurationEvent}; 
and the \texttt{Communicator}s,
which are used to establish the components' intra-communication.
%
To start, the \texttt{PulverisationRuntime} needs to know the name of the \emph{logical device},
the \emph{host} on which it is running, and the configurations produced by the two \ac{dsl}s.
%
At the start of the system, using the provided configuration, the runtime
spawns all the components that must execute on the \emph{deployment unit}.
%
Similarly, the \texttt{UnitReconfigurator} starts observing all the configured \texttt{ReconfigurationEvent}s
and registers the listeners for incoming reconfigurations.
%
Then,
each spawned component starts
executing its logic and sharing data with other components via the \texttt{Communicator},
which is responsible for the components' intra-communication by leveraging the communication protocol.
%
Finally,
when a reconfiguration occurs, the \texttt{UnitReconfigurator} considers the new deployment configuration and,
by interacting with the \texttt{Spawner},
starts or stops components in order to achieve the new setup.
